---
title: "The Last SLOme Stretch"
author: "Daphne Virlar"
date: "11/27/2018"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Daphne you are an angel sent from heaven 
```


```{r, packages}
#load data and clean it up
## install.packages("extrafont")  ## allows us to use more fonts than just the default
library(tidyverse) #for all things tidy
library(stargazer) 
library(extrafont) 
library(car)
#font_import() ## imports new fonts, need to give consent in console (y/n). May take a few minutes... but we now have all of the fonts to choose from!!!  


grad <- read_csv("grad_enrollment.csv")

grad <- grad %>% 
  select(Year, Total_Enrollment, Total_Males, Total_Females, Public_Enrollment, Private_Enrollment)


## Tidied PhD field data. I have removed most data in Excel, leaving only what we need to complete Part 2. 
phd_field <- read_csv("PhD_field.csv")

```



Part One: Compare trends in total graduate enrollment for males and females (including full-time/part-time and private/public universities) in the United States from 1967 - 2015.
```{r, part 1}
#performing linear regression to find trends in the data between M/F enrollment throughout time

female_lm <- lm(Total_Females ~ Year, data = grad)
male_lm   <- lm(Total_Males ~ Year, data = grad)

#Use stargazer to create regression table
#stargazer(female_lm, male_lm, type = "html")

stargazer(female_lm, male_lm, type = 'html', out="test.htm")
    #Comparing trends in male and female graduate enrollment by means of a linear model, 1967-2015

## is this table working for you?

summary(female_lm) 
summary(male_lm) 
  #The R^2 value tells us how much variance in our dependent variable is expained by our independent      variable
  #We use Multiple R-squared value for single linear regressions, and adjusted for multiple.


#Check for regression assumptions
plot(female_lm)
  #Uhm... what do we do when our model breaks assumptions of homoskedasticity?
      #Residuals vs Fitted shows our residuals violate the assumption of homoskedasticity. What does that mean for our R2 value? 

plot(male_lm)
  #Our data still violates homoskedasticity, but now are they also not normally distributed?? (QQ Plot)


#Use the cor.test() function to find Pearson’s r for the linear relationship described between year and graduate enrollment.
    #cor.test(x,y)
female_pr <- cor.test(grad$Year, grad$Total_Females)
female_pr

male_pr <- cor.test(grad$Year, grad$Total_Males)
male_pr


#Simplified -- Female
#The model (female graduate enrollment = 30,126*(year) - 58955502; no. of students) explains a significant amount of variance in female graduate enrollment (F(1,47) = 2669, p < 0.01, R2 = 0.98, Pearson's r = 0.99)

#Thorough -- Female
#Total female graduate enrollment is significantly predicted by year (b = -58955502, t(47) = 2669) with a strong positive correlation (Pearson's R = 0.99). The overall model (female enrollment = 30,126*year - 58955502, count of students) explains a significant amount of variance in total female graduate enrollment (F(1,47) = 2669, p < 0.001, R2 = 0.98).

#Simplified -- Male
#The model (male graduate enrollment = 9,069*(year) - 17112153; no. of students) explains a significant amount of variance in total male graduate enrollment (F(1,47) = 276, p < 0.001, R2 = 0.85, Pearson's r = 0.92)

##Thorough -- Male
##Total male graduate enrollment is significantly predicted by year (b = - 17112153, t(47) = Daphne where are you finding the female t-val? ) with a strong positive correlation (Pearson's R = 0.86). The model (male graduate enrollment = 9,069*(year) - 17112153; no. of students) explains a significant amount of variance in total male graduate enrollment (F(1,47) = 276, p < 0.001, R2 = 0.85, Pearson's r = 0.92)



######################################################################################################
#Explore the data

expo_graph <- ggplot(grad, aes(x = Year, y = Total_Males))+
  geom_point(aes(col = "Males"))+
  geom_point(aes(x = Year, y = Total_Females, col = "Females"))

#Plot the data  
enroll_graph <- ggplot(grad, aes(x = Year, y = Total_Males))+
  geom_point(aes(col="Male"))+
  geom_smooth(method = lm, se = F, aes(col = "Male"))+
  geom_smooth(aes(x = Year, y = Total_Females, col = "Female"), method = lm, se = F)+
  geom_point(aes(x = Year, y = Total_Females, col = "Female"))+
  ylab("Enrollment \n")+
  xlab("\n Year")+
  #ggtitle("Changes in University Enrollment History by Sex, 1967-2015 \n")+
  theme_classic()

enroll_graph + theme(plot.title = element_text(hjust = 0.50),
                  text = element_text(size = 12, 
                                      family ="Times New Roman"))+  ## font shoes in knitted document
  scale_color_discrete(name = "Sex")

```




Part Two. Shifts in female PhD recipients by field (1985, 2000, and 2015).
```{r, part 2}
#Describe if and how there was a shift in PhDs awarded to females in four fields in 1985, 2000, and 2015

#filter out male from the data set
phd_filter <- phd_field %>% 
  filter(sex == "female") %>%  #we're only keeping females in the data set. fuck the men. get rid of them
  select(field, year, number, percent) %>% 
  #I was having trouble changing the legend status', so I just changed the actual row names.
   mutate(field = case_when(
    field == "physical and earth sciences" ~ "Physical Sciences",
    field == "Engineering" ~ "Engineering",
    field == "Education" ~ "Education",
    field == "Humanities and arts" ~ "Humanities"))


#after trying to do exploratory graphs, i quit because they didn't show anything useful. That's my ted talk, thanks for coming. However, because of the Central Limit Theorem, we can assume things are normally distributed. 

## Totally fine!! 


#Question being asked: What does the trend look like for female's receiving phds over the years, by field?
    #graph should have data points for each field, and x axis should be years
phd_graph <- ggplot(phd_filter, aes(x = year, y = number, group = field))+
  geom_line(aes(col = field))+
  geom_point(aes(col = field))+
  ylab("Degrees Awarded \n")+ #Need to change this? 
  xlab("\nYear")+
  #ggtitle("Total Doctorate Degrees Awarded by Field, 1985-2015 \n")+
  theme_classic()+
  scale_x_continuous(breaks = c(1985, 2000, 2015))
  

phd_graph + theme(plot.title = element_text(hjust = 0.50),
                  text = element_text(size = 12, 
                                      family ="Times New Roman")) + #to center the title, and change fon't size. This will be helpful in the actual document we turn in. 
  scale_color_discrete(name = "Field of Study") #This the legend title








######################################################################################################

#Statistical analysis for part 2

#Thought process: Using chi-square to test for independence between doctorates and fields of study
#Lecture 12

          #In General
#H0: No association between variables studied. The null hypothesis is independence.
#HA: There IS an association between variables.

          #For Us
#H0: There is no association between doctorates awarded and field of study
#HA: There IS an associaiton between 



#Create new data frame for use in Chi-Square
Year85 <- c(3491, 198, 1392, 569)
Year00 <- c(4179, 838, 2672, 1022)
Year15 <- c(3502, 2301, 2832, 1988)

chi_df <- rbind(Year85, Year00, Year15)
colnames(chi_df) <- c("Education", "Engineering", "Humanities", "Physical Sciences")

chi_test <- chisq.test(chi_df)

chi_test
  #X-squared = 2073.3, df = 6, p-value < 2.2e-16
  #We reject the null hypothesis! Degrees awarded ARE associated with field of study!

## for some reason Natalie's chi^2 is not running, so I can't proof-read this. 

#"There is a significant association between time elapsed and degrees awarded by department (X2(6) = 2073.3, p < 0.001)"

#Degrees awarded by field (education, engineering, humanities, physical sciences) differed significantly between 1985, 2000, and 2015 (X2(6) = 2073.3, p < 0.001).

```








Part 3. Compare median salaries for male and female doctorate recipients in 2015. Answer these two questions: 

(1) Does median salary differ significantly between male and female starting postdoc positions? 

(2) Does median salary differ significantly between male and female PhD recipients in non-postdoc employment positions?
```{r}
#Look at lecture 11. To compare medians, we use Mann-Whitney U. HOWEVER. Because the data is paired, we use Wilcoxon Signed-Rank (Lab 7)
#Male and female salaries for starting postdoctoral and other employment positions (2015)

#load in the data
phd_salary <- read_csv("PhD_Salary.csv")

postdoc <- wilcox.test(phd_salary$Male_Postdoc, phd_salary$Female_Postdoc, paired = TRUE)
non_postdoc <- wilcox.test(phd_salary$Male_Employment, phd_salary$Female_Employment, paired = TRUE)

#H0: Ranks are equal
#HA: Ranks are NOT equal
  #Samples are not independent
  #Data is paired because we are comparing 


postdoc

##data:  phd_salary$Male_Postdoc and phd_salary$Female_Postdoc
##V = 19.5, p-value = 0.8884

##Retain the null

non_postdoc

##data:  phd_salary$Male_Employment and phd_salary$Female_Employment
##V = 101, p-value = 0.002572

##Accept the alternative

```



Part 4. Exploring academic salaries for professors in U.S. colleges.
```{r}
#Develop a model describing faculty salary based on data for faculty sex, rank, years in current position, field, and number of years since doctoral degree was earned.

faculty <- read_csv("faculty_salary.csv")

faculty$rank <- fct_relevel(faculty$rank, "AsstProf")
  #This takes that level as the reference level
  levels(faculty$rank)


fac_lm1 <- lm(salary ~ sex + years_service + discipline + postgrad_years + rank, data = faculty)
fac_lm2 <- lm(salary ~ sex + discipline + rank, data = faculty)


summary(fac_lm1)
summary(fac_lm2)

#Figure out which x variables are eing compounded with each other, and take them out of the model. 
vif(fac_lm2)
  #Can get rid of years_service and post grad years since they're collinear. Explain that they are better off explained in rank (Prof, Ass. Prof, Asst. Prof)




#Compare models using AIC
    #Recall: lower AIC value indicates a *more optimal* balance - **BUT STATISTICS IS NO SUBSTITUTE FOR JUDGEMENT!!!**
fac1_aic <- AIC(fac_lm1)
fac2_aic <- AIC(fac_lm2)

fac1_aic; fac2_aic
# 9093     9095 
    #Find a resource that has actually studied this, and cite it in the paper. Most likely, we will end up actually using the second linear model because despite having a slightly higher AIC, it doesn't include the counterintuitive negative coefficient associated with length of service



#Create stargazer table 
stargazer(fac_lm2, fac_lm2, type = 'html', out="AIC.htm")


```

